### 2018-04-08 ident/docs/draft_2.tex
changes to methods, results and figure 1

	The development of these enzyme kinetic rate laws is based on in vitro observations of enzyme activity. Accordingly, authors have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions, as opposed to in vitro conditions \parencite{Heijnen2005,Heijnen2013}. This problem is further compounded by the fact that typical parameter values used in kinetic models of metabolism are estimated in vitro, and may not be applicable under in vivo conditions \parencite{Heijnen2005,Smallbone2007}. Also, parameters estimated based on in vivo experimental data are associated with large uncertainties \parencite{Link2014}. Using in vitro, or unreliable in vivo parameter estimates to study the steady state and dynamic responses of metabolic networks, under in vivo conditions, to different perturbations reduces confidence in the model predicted behaviour. Consequently, this hampers the use of these models to gain insight into metabolic network function \parencite{Andreozzi2016a,Vasilakou2016}, and the increase in prediction uncertainty becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design \parencite{Saa2017}.

@@

	In spite of the development of these methods to quantify parameter estimation uncertainty, model parameter identifiability, a necessary, and sometimes sufficient condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}. Briefly, it concerns with the ability to estimate unique values for all model parameters from observed experimental data. In a model, any parameter is said to be structurally or a priori identifiable if its values can be estimated independent of all other model parameters from available experimental data. However, if parameters cannot be estimated independent of each other due to redundant model parameterization, or due to the nonlinear relationship between the model parameters, then the parameters are said to be structurally non-identifiable. Conversely, if the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, and the uncertainty in parameter estimates is unquantifiable, the parameter is said to be practically non-identifiable \parencite{Ljung1994}.

@@

	The parameters in $\theta$ in any nonlinear model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \left[x, v\right]^T = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
	\begin{align}\label{eq:stident}
	\Phi(\theta_1,u) = \Phi(\theta_2,u) \iff \theta_1 = \theta_2
	\end{align}
	Accordingly, if parameters in $\theta$ have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, they are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively. So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unmeasured and measured concentrations/fluxes, as well as correlations between different model parameters \textcolor{red}{Rudiyanto Gunawan paper on model discrimination}. Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.

@@

	So, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated from measurable data. On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.

@@

	In order to establish the practical identifiability of kinetic parameters for each flux $v_i$, $i={1, 2, ..., n_r}$, we describe a computer algebra-based method. The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\mathbf{\theta}$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b). This is done by first selecting a combination of $p$ data sets from the $n_{E}$ available experimental data sets. The combination consisting of data from $p$ different experiments is then used to formulate a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}). 
	\begin{align}\label{eq:nonlineq}
	v_{i, j} = f_j(\mathbf{x}_j,\mathbf{\theta}, \mathbf{u}_j) && \forall j=\{1, 2, ..., p\}\subset\{1, 2, ..., n_E\}
	\end{align}
	Here, $v_{i,j}$ refers to the value of the flux $v_i$ obtained from experiment $j$. $\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$, whose elements are denoted by $\theta_k$.

	Each equation in (\ref{eq:nonlineq}), indicated by the index $j$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for each $v_i$, $i={1, 2, ..., n_r}$, described in Equation (\ref{eq:kinstoich}b), written for concentrations ($\mathbf{x}_j$, $\mathbf{u}_j$) and fluxes ($v_{i,j}$) obtained from experiment $j$. Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for each parameter $\theta_k$ in $\theta \in \mathbb{R}^p$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b). Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from all chosen experiments $j = {1, 2, ..., p}$, in the selected combination.
	\begin{align}\label{eq:theta-eq}
	\theta_k = g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = \frac{N_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}{D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}
	\end{align}

@@

	\subsection{Kinetic model of gluconeogenesis in \textit{E. coli}}\label{sec:small-model}
	A previously proposed kinetic model \parencite{Kotte2014, Srinivasan2017} for acetate consumption through gluconeogenesis (Figure \ref{fig:network}) is used as a case study to illustrate the utility of identifiability analysis for experimental design for parameter estimation in kinetic models of metabolism. The kinetic model is described below.

	\begin{equation}\label{eq:ode1}
	\frac{d}{dt}pep=v_1-v_2-v_4
	\end{equation}
	\begin{equation}\label{eq:ode2}
	\frac{d}{dt}fdp=v_2-v_3
	\end{equation}
	\begin{equation}\label{eq:ode3}
	\frac{d}{dt}E=v_5 - d E
	\end{equation}
	The kinetic expressions for fluxes $v_1$ through $v_5$ are given below. The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics. The acetate flux through $v_1$ is also governed by the quantity of available enzyme E. 
	\begin{equation}\label{eq:flux1}
	v_1 = k_{1}^{cat}E\frac{ac}{ac+K_{1}^{ac}}
	\end{equation}	
	\begin{equation}\label{eq:flux2}
	v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
	\end{equation}
	\begin{equation}\label{eq:flux3}
	v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
	\end{equation}
	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.

@@

	\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis}\label{sec:experiments}
	Steady state information on the metabolome and the fluxome can be gathered under different physiological conditions. One way to alter the physiological conditions is to change the substrate concentration under which the cells grow. In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$. Thus, steady state metabolite concentrations and fluxes can be calculated under various acetate concentrations to form different experimental data combinations that measure cellular response to changes in the substrate concentration.

	Physiological changes can also be brought about by perturbing the expression levels for different enzymes within a metabolic network. The model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model} has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes. Accordingly, in addition to measuring network response to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations. In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico. The 21 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.

@@

	Similarly, for identifying $v_3$ that is described by three parameters ($\theta \in \mathbb{R}^3$), we need data from a combination of three different experiments. Based on the 21 experiments in Table \ref{tab:pval}, we have 7980 distinct combinations (21 x 20 x 19) of three experiments. However, it is important to note that not all of the either 420 combinations (for $v_1$ and $v_2$), or the 7980 combinations (for $v_3$), may be suitable for parameter estimation. Hence, the actual number of data combinations that can be used for practically identifying parameters in these fluxes would be smaller than the aforementioned numbers. Moreover, among these chosen smaller number of experimental data combinations, our methodology enables us to determine the number and nature of experimental data combinations, and consequently, the experiments that might allow for parameter identification, and subsequent parameter estimation. The following section describes a procedure that can be used to design experiments within this context.

@@

	However, not all metabolite concentrations and fluxes in the model (Equation \ref{eq:kinstoich}) change for any random experiment. This makes unambiguous estimation of parameters impossible, either due to the inherent correlation between changes in different concentrations or fluxes, or due to the homeostasis of the concentrations and fluxes under the chosen experimental conditions \parencite{Heijnen2013}. In such scenarios, the need to design experiments to effect a change in, and discriminate between changes in different concentrations/fluxes becomes necessary.

@@

	The identifiability of each parameter based on each experiment indexed as $j = \{1, 2, ..., n_E\}$ is established based on the methodology described in section \ref{sec:ident} (Figure \ref{fig:ident-flowchart}b), and demonstrated in section \ref{sec:example}. Subsequently, for any flux $v_i$, and for any combination of $p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., p}$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable. Consequently, the corresponding experiment cannot be used for estimating parameter $\theta_k$, and needs to be discarded from the set of all necessary experiments. Furthermore, another experiment from $j = \{1, ..., n_E\}$ needs to be selected to replace the discarded experiment such that parameter $\theta_k$ is identifiable. This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$. In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$. Note that if none of the $n_E$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n_E+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability. This analysis can be performed for each flux in a metabolic network independent of all the other fluxes, making it theoretically scalable even to genome-scale models of metabolism.

@@

	\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis}\label{sec:example}	
	The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}). Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression previously given for $v_1$ (Equation \ref{eq:flux1}) cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$. So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}). Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$. The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
	\begin{align}\label{eq:flux1a}
	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
	\end{align}		
	We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability.

	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ still holds for $v_1$, as mentioned in section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments. So, from the $n_E = 21$ different experiments described in section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$. We label the available concentrations and fluxes as ${ac}^{(j)}$ and ${v_1}^{(j)}$, respectively. Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
	\begin{align*}%\label{eq:nonlin-flux2}
	{v_1}^{(j)} = V_{1}^{max}\frac{ac^{(j)}}{ac^{(j)}+K_{1}^{ac}(ne)} &&  j=\{1, 2\}
	\end{align*}

	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b). These expressions have the form shown in Equation (\ref{eq:theta-eq}).
	\begin{subequations}\label{eq:v1_par}
		\begin{align}		
		\theta_1 &= V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
		\theta_2 &= K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
		\end{align}
	\end{subequations}
	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations. Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).

@@

	\subsection{Maximum reaction rates are more identifiable than enzyme binding constants}\label{sec:trends}	
	In Figure \ref{fig:figure1} we show the number and percentage of combinations that are capable of identifying each parameter in each flux. Based on the definition given in section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter. The three panels in Figure \ref{fig:figure1} represent the degree of identifiability for parameters modeling the three different fluxes of the small network individually.

@@

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1/v1_k1cat_v1max_v2_v3_1_2_ident}}
		\caption{The number of data combination from 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data. The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter. Since the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, the total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower (at 306) that the total number of combinations used to test identifiability of parameters for $v_3$ (at 4896), which is modeled with three parameters and requires data from at least three different experiments.}\label{fig:ident}
	\end{figure}

>>>>

	The development of these enzyme kinetic rate laws is based on in vitro observations of enzyme activity. Accordingly, authors have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions, as opposed to in vitro conditions \parencite{Heijnen2005,Heijnen2013}. This problem is further compounded by the fact that typical parameter values used in kinetic models of metabolism are estimated in vitro, and may not be applicable under in vivo conditions \parencite{Heijnen2005,Smallbone2007}. Also, parameters estimated based on in vivo experimental data are associated with large uncertainties \parencite{Link2014}. Using in vitro, or unreliable in vivo parameter estimates to study the steady state and dynamic responses of metabolic networks to different perturbations, under in vivo conditions, reduces confidence in the model predicted behaviour. Consequently, this hampers the use of these models to gain insight into metabolic network function \parencite{Andreozzi2016a,Vasilakou2016}, and the increase in prediction uncertainty becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design \parencite{Saa2017}.

@@

	In spite of the development of these methods to quantify parameter estimation uncertainty, model parameter identifiability, a necessary, and sometimes sufficient condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}. Briefly, it concerns with the ability to estimate unique values for all model parameters from observed experimental data. In a model, any parameter is said to be structurally or a priori identifiable if its values can be uniquely estimated independent of all other model parameters from available experimental data. However, if parameters cannot be uniquely estimated independent of each other due to redundant model parameterization, or due to the nonlinear relationship between the model parameters, then the parameters are said to be structurally non-identifiable. Conversely, if the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, and the uncertainty in parameter estimates is unquantifiable, the parameter is said to be practically non-identifiable \parencite{Ljung1994}.

@@

	The parameters in $\theta$ in any nonlinear model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \left[x, v\right]^T = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
	\begin{align}\label{eq:stident}
	\Phi(\theta_1,u) = \Phi(\theta_2,u) \iff \theta_1 = \theta_2
	\end{align}
	Accordingly, if parameters in $\theta$ have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, they are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively. So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unmeasured and measured concentrations/fluxes, as well as correlations between different model parameters \textcolor{red}{(Rudiyanto Gunawan paper on model discrimination)}. Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.

@@

	So, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but are also parameterized in such a way that the value of these parameters can be estimated from measurable data. On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.

@@

	In order to establish the practical identifiability of kinetic parameters for each flux $v_i$, $i={1, 2, ..., n_r}$, we describe a computer algebra-based method. The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\mathbf{\theta}$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b). This is done by first selecting a combination of $p\le n_E$ experimental data. The fluxes and concentrations from $p$ different experiments are then used to formulate a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}). 
	\begin{align}\label{eq:nonlineq}
	v_{i, j} = f_j(\mathbf{x}_j,\mathbf{\theta}, \mathbf{u}_j) && \forall j=\{1, 2, ..., p\}\subset\{1, 2, ..., n_E\}
	\end{align}
	Here, $v_{i,j}$ refers to the value of the flux $v_i$ obtained from experiment $j$. $\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$, whose elements are denoted by $\theta_k$.

	Each equation in (\ref{eq:nonlineq}), indicated by the index $j$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for each $v_i$, $i={1, 2, ..., n_r}$, described in Equation (\ref{eq:kinstoich}b), written for concentrations ($\mathbf{x}_j$, $\mathbf{u}_j$) and fluxes ($v_{i,j}$) obtained from experiment $j$. Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for each parameter $\theta_k$ in $\theta \in \mathbb{R}^p$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b). Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from $p$ experiments denoted by the index $j = {1, 2, ..., p}$.
	\begin{align}\label{eq:theta-eq}
	\theta_k = g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = \frac{N_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}{D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}
	\end{align}

@@

	\subsection{Kinetic model of gluconeogenesis in \textit{E. coli}}\label{sec:small-model}
	A previously proposed kinetic model \parencite{Kotte2014, Srinivasan2017} for acetate consumption through gluconeogenesis (Figure \ref{fig:network}) is used as a case study to illustrate identifiability analysis for experimental design for parameter estimation in kinetic models of metabolism. The kinetic model is described below.

	\begin{equation}\label{eq:ode1}
	\frac{d}{dt}pep=v_1-v_2-v_4
	\end{equation}
	\begin{equation}\label{eq:ode2}
	\frac{d}{dt}fdp=v_2-v_3
	\end{equation}
	\begin{equation}\label{eq:ode3}
	\frac{d}{dt}E=v_5 - d E
	\end{equation}
	The kinetic expressions for fluxes $v_1$ through $v_5$ are given below. The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics. The acetate flux through $v_1$ is also governed by the quantity of available enzyme E. 
	\begin{equation}\label{eq:flux1}
	v_1 = k_{1}^{cat}E\frac{ac}{ac+K_{1}^{ac}}
	\end{equation}		
	The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}). Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression in Equation (\ref{eq:flux1}) for $v_1$ cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$. So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}). Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$. The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
	\begin{align}\label{eq:flux1a}
	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
	\end{align}		
	We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability. 	
	\begin{equation}\label{eq:flux2}
	v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
	\end{equation}
	\begin{equation}\label{eq:flux3}
	v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
	\end{equation}
	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.

@@

	\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis}\label{sec:experiments}
	Steady state information on the metabolome and the fluxome can be gathered under different physiological conditions. One way to alter the physiological conditions is to change the substrate concentration under which the cells grow. In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$ (Equation \ref{eq:flux1}). Thus, steady state metabolite concentrations and fluxes can be calculated under various acetate concentrations to form different experimental data combinations that measure cellular response to changes in the substrate concentration.

	Physiological changes can also be brought about by perturbing the expression levels for different enzymes within a metabolic network. The model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model} has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes. Accordingly, in addition to measuring network responses to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations. In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico. The 21 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.

@@

	Similarly, for identifying $v_3$ that is described by three parameters ($\theta \in \mathbb{R}^3$), we need data from a combination of three different experiments. Based on the 21 experiments in Table \ref{tab:pval}, we have 7980 distinct combinations (21 x 20 x 19) of three experiments. However, it is important to note that not all of the either 420 combinations (for $v_1$ and $v_2$), or the 7980 combinations (for $v_3$), may be suitable for parameter estimation. Hence, the actual number of data combinations that can be used for practically identifying parameters in these fluxes would be smaller than the aforementioned numbers. Moreover, among the chosen smaller number of experimental data combinations, the methodology that we propose enables us to determine the number and nature of experimental data combinations, and consequently, the experiments that might allow for parameter identification, and subsequent parameter estimation. The following section describes a procedure that can be used to design experiments within this context.

@@

	Not all metabolite concentrations and fluxes in the model (Equation \ref{eq:kinstoich}) change for any random experiment. This makes unambiguous estimation of parameters impossible, either due to the inherent correlation between changes in different concentrations or fluxes, or due to the homeostasis of the concentrations and fluxes under the chosen experimental conditions \parencite{Heijnen2013}. In such scenarios, the need to design experiments to effect a change in, and discriminate between changes in different concentrations/fluxes becomes necessary.

@@

	The identifiability of each parameter based on each experiment with index $j = {1, 2, ..., n_E}$ is established based on the methodology described in section \ref{sec:ident} (Figure \ref{fig:ident-flowchart}b), and demonstrated in section \ref{sec:example}. Subsequently, for any flux $v_i$, and for any combination of $p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., p}$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$ (Figure \ref{fig:ident-flowchart}b), then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable. Consequently, the corresponding experiment cannot be used for estimating parameter $\theta_k$, and needs to be discarded from the set of all necessary experiments. Furthermore, another experiment from $j = {1, ..., n_E}$ needs to be selected to replace the discarded experiment such that parameter $\theta_k$ is identifiable. This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$. In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$. Note that if none of the $n_E$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n_E+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability. This analysis can be performed for each flux in a metabolic network independent of all the other fluxes, making it theoretically scalable even to genome-scale models of metabolism.

@@

	\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis}\label{sec:example}	
	In this section, we illustrate the methodology that we developed to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network}) described in section \ref{sec:small-model}. We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) for this demonstration.

	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments. So, from the $n_E = 21$ different experiments described in section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$. We label the available concentrations and fluxes as ${ac}^{(j)}$ and ${v_1}^{(j)}$, respectively. Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
	\begin{align*}%\label{eq:nonlin-flux2}
	{v_1}^{(j)} = V_{1}^{max}\frac{ac^{(j)}}{ac^{(j)}+K_{1}^{ac}(ne)} &&  j=\{1, 2\}
	\end{align*}

	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b). These expressions have the form shown in Equation (\ref{eq:theta-eq}).
	\begin{subequations}\label{eq:v1_par}
		\begin{align}		
		\theta_1 &= V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
		\theta_2 &= K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
		\end{align}
	\end{subequations}
	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations. Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b). We apply the method described in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}, and illustrated above for $v_1$ to fluxes $v_2$ and $v_3$ in Figure \ref{fig:network}. We summarize the results on the identifiability of parameters in these fluxes in Figure \ref{fig:ident} and describe them below.

	In Figure \ref{fig:ident} we show the number and percentage of combinations that are capable of identifying each parameter in each flux. Based on the definition given in section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter. The three panels in Figure \ref{fig:ident} represent the degree of identifiability for parameters modeling the three different fluxes of the small network individually.

	\subsection{Maximum reaction rates are more identifiable than enzyme binding constants}\label{sec:trends}

@@

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1/v1_v2_v3_v5_ident}}
		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$ \textcolor{red}{(need to change figure)} when there is no noise in the input experimental data. The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter. The total number of experiments used to identify parameters in $v_1$, $v_2$ and $v_5$ is two. The total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is 240, 421 different combinations are used to test identifiability for $v_5$, and 4896 combinations are used to test identifiability of parameters for $v_3$. Note that $v_3$ is modeled with three parameters and requires data from at least three different experiments, and also has two possible roots for each of its parameters. Results for only one of the two roots indicated by (1) is shown here.}\label{fig:ident}
	\end{figure}

@@
